How to \emph{not} write Virtual Machines for Dynamic Languages

### 2007-05-27 talk/dyla2007/dyla.tex
Small fixes.

In this paper, we will argue that it is possible to
benefit from and integrate with OO VMs while keeping the dynamic
language implemented with a single, simple source code base.  The idea is
to write an interpreter for that language in another sufficiently
high-level but less dynamic language.  This interpreter plays the role
of a specification for the dynamic language.  With a sufficiently capable
translation toolchain we can then generate whole virtual machines from
this specification -- either wholly custom VMs for C-level operating
systems or VMs layered on top of various OO VMs.  In other words,
meta-programming techniques can be used to successfully replace a
foreseeable one-VM-fits-all standardization attempt.

@@

In the sequel (??? - what is the sequel), we will focus on the relative advantages and
inconveniences of the PyPy approach compared to the approach of
hand-writing a language implementation on top of an OO VM.

@@

Our approach -- a single ``meta-written'' implementation -- naturally
leads to language implementations that have various advantages over the
``hand-written'' implementations.  Firstly, it is a single-source
approach -- we explicitly seek to solve the problem of proliferation of
implementations.  In the sequel (???), we will show that this goal can be
achieved without giving up the advantages of hand-written
implementations for OO VMs.  Moreover, there are additional advantages which,
in our opinion, are significant enough to hint that meta-programming,
though not widely used in general-purpose programming, is an essential
tool in a language implementer's toolbox.

>>>>

In this paper, we will argue that it is possible to
benefit from and integrate with OO VMs while keeping the dynamic
language implemented with a single, simple source code base.  The idea is
to write an interpreter for that language in another sufficiently
high-level but less dynamic language.  This interpreter plays the role
of a specification for the dynamic language.  With a sufficiently capable
translation toolchain we can then generate whole virtual machines from
this specification -- either wholly custom VMs for C-level operating
systems or as a layer on top of various OO VMs.  In other words,
meta-programming techniques can be used to successfully replace a
foreseeable one-VM-fits-all standardization attempt.

@@

In the sequel, we will focus on the relative advantages and
inconveniences of the PyPy approach compared to the approach of
hand-writing a language implementation on top of an OO VM.

@@

Our approach -- a single ``meta-written'' implementation -- naturally
leads to language implementations that have various advantages over the
``hand-written'' implementations.  Firstly, it is a single-source
approach -- we explicitly seek to solve the problem of proliferation of
implementations.  In the sequel, we will show that this goal can be
achieved without giving up the advantages of hand-written
implementations for OO VMs.  Moreover, there are additional advantages which,
in our opinion, are significant enough to hint that meta-programming,
though not widely used in general-purpose programming, is an essential
tool in a language implementer's toolbox.
